\documentclass{article}
\input{macros}

% \usepackage{fullpage}

\begin{document}
\title{Adjoint Logic over a 2-Category of Worlds}
\author{Jason Reed}
\maketitle
\def\rok{\mathop{\mathrm{ok}}\nolimits}
\def\lol{\multimap}
\def\trov{\triangleright}

\section{Syntax}
Fix a 2-category of worlds (or ``judgmental levels'') with objects $p,q,\ldots$, morphisms $\alpha,\beta,\ldots$, and 2-cells $e,\ldots$. We're going to give a proof theory of the language
$$\begin{tabular}{rl@{\ }l@{\ \ }l}
Positive Propositions &$P$&$::=$&$ F_\alpha P \celse \dns N \celse a^+ \celse  P \tensor P \celse 1 \celse P \oplus P \celse 0 $\\
Negative Propositions &$N$&$::=$&$ U_\alpha N \celse \ups P \celse a^- \celse  P \lol N \celse N \amp N \celse \top $\\
\end{tabular}$$
first via translation into focused first-order logic, then directly by giving proof rules.

\section{World-typing}
The translation needs some extra syntax to describe the first-order individuals involved.
$$\begin{tabular}{rl@{\ }l@{\ \ }l}
Resources&$r$&$::=$&$ F_\alpha r\celse \rho \celse r \tensor r \celse 1$\\
Frames&$f$&$::=$&$ U_\alpha f\celse \phi\celse r \lol f $\\
Structures&$s$&$::=$&$ r \trov f $\\
\end{tabular}$$
The subscript $\alpha$ ranges over all morphisms in the 2-category.
The symbols $\rho$ and $\phi$ are metavariables for resource variables
and frame variables respectively. Every resource and frame is `located
at' a `world', i.e. an object in the underlying 2-category of worlds.
We can make a typing judgment to capture the appropriate invariant.
$\Delta$ is a context of assumptions $\rho : p$ and $\phi : p$ about
resource and frame variables. The intuition behind $\Delta \prov r :
p$ and $\Delta \prov f : p$ is that $r$ and $f$ are like a zipper with
a filler, together describing $r \trov f$, the structure of the
sequent. The invariant is just that the `type' of the zipper's hole,
the object of the 2-category at which we've split it in half, must
match the data filling that hole.
$$\begin{prooftree}
  \justifies
  \Delta, \rho : p \prov \rho : p
\end{prooftree}
\qquad
\begin{prooftree}
  \justifies
  \Delta, \phi : p \prov \phi : p
\end{prooftree}
$$
$$\begin{prooftree}
  \Delta \prov r : p\qquad \alpha : p \to q
  \justifies
  \Delta \prov F_\alpha r : q
\end{prooftree}
\qquad
\begin{prooftree}
  \Delta \prov f : q\qquad \alpha : p \to q
  \justifies
  \Delta \prov U_\alpha f : p
\end{prooftree}
$$
$$\begin{prooftree}
  \Delta \prov r_1:p \qquad\Delta \prov r_2 : p
  \justifies
  \Delta \prov r_1 \tensor r_2: p
\end{prooftree}
\qquad
\begin{prooftree}
  \justifies
  \Delta \prov 1: p
\end{prooftree}
\qquad
\begin{prooftree}
  \Delta \prov r:p \qquad\Delta \prov f : p
  \justifies
  \Delta \prov r \lol f: p
\end{prooftree}
$$
$$
\begin{prooftree}
  \Delta \prov r:p \qquad\Delta \prov f : p
  \justifies
  \Delta \prov r \trov f \rok
\end{prooftree}
$$ It may look like resources and frames recapitulate most of the
syntax of propositions, but it's critical that they lack shifts (less
critical, but true, is the fact that they lack additives).

\section{Translation}
The target language is focused first order intuitionistic logic,
with two fixed atomic propositions $b^+$ and $\#^-$. The
translation consists of two functions: $N^f$ takes a negative
proposition and a frame, and $P^r$ takes a positive proposition and a
resource, and both yield a target-language positive proposition,
regardless of the polarity of the input. They are defined in
Figure~\ref{figure.translation}.



\begin{figure}
$$\begin{tabular}{ccc}
\toprule
$N$&$ N^f$\\
\midrule
$U_\alpha N$&$ \exists \phi . N^\phi  \land f = U_\alpha \phi $\\
$P \lol N$&$ \exists \rho . P^\rho \land \exists \phi . N^\phi \land f = \rho \lol \phi $\\
$N_1 \amp N_2$&$N_1^f\lor N_2^f$\\
$\top$&$0$\\
$\ups P$&$ \dns \forall\rho .   P^\rho \imp \#^-(\rho \trov f) $\\
$ a^-$&$b^+(a^-, f) $\\
\bottomrule
\end{tabular}$$

$$\begin{tabular}{ccc}
\toprule
$P$&$ P^r$\\
\midrule
$F_\alpha P$&$ \exists \rho . P^\rho  \land r = F_\alpha \rho $\\
$P_1 \tensor P_2$&$ \exists \rho_1 . P_1^{\rho_1} \land  \exists \rho_2 . P_2^{\rho_2}\land r = \rho_1 \tensor \rho_2$\\
$P_1 \oplus P_2$&$P_1^r\lor P_2^r$\\
$1$&$r = 1$\\
$0$&$0$\\
$\dns N$&$ \dns \forall\phi .   N^\phi \imp \#^-(r \trov \phi) $\\
$ a^+$&$b^+(a^+, r) $\\
\bottomrule
\end{tabular}$$
\caption{Translation}
\label{figure.translation}
\end{figure}




\input{foc-rules}






\end{document}
